Eco-evolutionary dynamics of finite populations from first principles

### 2023-04-04 ind_files/99_Appendix_D_examples.tex
Correct minor typos

For this model, we can in fact quantitatively derive the effects of noise-induced selection by explicitly calculating each term of equation \eqref{nD_stochastic_RM}. Let $p = x_1/(x_1 + x_2)$ be the frequency of type 1 individuals and let $q = 1-p$. Then, it is easy to check by direct substitution of our functional forms that \eqref{nD_stochastic_RM} becomes
\begin{equation}
	dp =  \frac{2}{K}p^2qdt + \mu\left(1-\frac{1}{KN_K}\right)\left(q-p\right) + q\sqrt{A^{+}_1}dW_t^{(1)} - p\sqrt{A^{+}_2}dW_t^{(2)}
\end{equation}
where $A_i^{+}$ is as defined throughout this thesis and each $W^{i}$ is an independent Wiener process. This equation clearly shows the biasing effect of noise-induced selection in the first term of the RHS. Since $p^2(1-p) > 0$ for $p \in (0,1)$, this term always tends to increase the fraction of type 1 individuals in the population. The second term is a mutational term, and simply reflects the fact that we assumed that $1 \to 2$ and $2 \to 1$ mutations are equally likely, and thus, the net effect of mutational effects depends on the difference between the frequencies of type 1 and type 2 individuals and flows towards the type with lower frequency, which is why it is $q-p$ and not $p-q$. Finally, the two $dW_t$ terms are non-directional and vanish upon taking an expectation over the probability space, and therefore have no net contribution other than `blurring out the results' if we look at the dynamics averaged over many realizations.

>>>>

For this model, we can in fact quantitatively derive the effects of noise-induced selection by explicitly calculating each term of equation \eqref{nD_stochastic_RM}. Let $p = x_1/(x_1 + x_2)$ be the frequency of type 1 individuals and let $q = 1-p$. Then, it is easy to check by direct substitution of our functional forms that equation \eqref{nD_stochastic_RM} becomes
\begin{equation}
	dp =  \left[\frac{2}{K}p^2q + \mu\left(1-\frac{1}{KN_K}\right)\left(q-p\right)\right]dt + q\sqrt{A^{+}_1}dW_t^{(1)} - p\sqrt{A^{+}_2}dW_t^{(2)}
\end{equation}
where $A_i^{+}$ is as defined throughout this thesis and each $W^{(i)}_t$ is an independent Wiener process. This equation clearly shows the biasing effect of noise-induced selection in the first term of the RHS. Since $p^2(1-p) > 0$ for $p \in (0,1)$, this term always tends to increase the fraction of type 1 individuals in the population. The second term is a mutational term, and simply reflects the fact that we assumed that $1 \to 2$ and $2 \to 1$ mutations are equally likely, and thus, the net effect of mutational effects depends on the difference between the frequencies of type 1 and type 2 individuals and flows towards the type with lower frequency, which is why it is $q-p$ and not $p-q$. Finally, the two $dW_t$ terms are non-directional and vanish upon taking an expectation over the probability space, and therefore have no net contribution other than `blurring out the results' if we look at the dynamics averaged over many realizations.
